Building a GNU/Hurd Cluster

### 2018-01-15 papers/cua.tex
cua lecture - working on memory management, and about to reorganize Tex IfSubStr

\usepackage{color}
\usepackage{comment}
\usepackage{graphicx}

@@

\begin{frame}
\frametitle{Cluster Computing - History}
\begin{itemize}
\item Beowulf
\item Hadoop
\item Cluster Linux
Kerrighed - Latest version is 3.0.0 which was released on 14 June 2010 and based on Linux kernel 2.6.30.
OpenSSI - last development release 2010; last stable release 2007
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Techniques to exploit parallelism}
\begin{itemize}
\item Instruction level parallelism
\begin{itemize}
\item Superscalar vector operations
\item Speculative execution
\end{itemize}
\item Thread level parallelism
\item Node level parallelism
\begin{itemize}
\item Clusters
\end{itemize}
\end{itemize}
\end{frame}

@@

int main () \{
  for (int i=0; i<512; i++) \{
    a[i] = b[i] + c[i];
  \}
\}

gcc -O2 -o ex2 ex2.c

00000000004003e0 <main>:
4003e0:31 c0                xor    %eax,%eax
4003e2:66 0f 1f 44 00 00    nopw   0x0(%rax,%rax,1)
4003e8:8b 90 60 18 60 00    mov    0x601860(%rax),%edx
4003ee:03 90 60 10 60 00    add    0x601060(%rax),%edx
4003f4:48 83 c0 04          add    $0x4,%rax
4003f8:89 90 5c 20 60 00    mov    %edx,0x60205c(%rax)
4003fe:48 3d 00 08 00 00    cmp    $0x800,%rax
400404:75 e2                jne    4003e8 <main+0x8>
400406:31 c0                xor    %eax,%eax
400408:c3                   retq

@@

\begin{frame}[fragile]
\frametitle{Adding 512 64-bit integers - SSE superscalar instructions}
\begin{semiverbatim}
\tiny
int a[512] __attribute__((aligned(0x10)));
int b[512] __attribute__((aligned(0x10)));
int c[512] __attribute__((aligned(0x10)));

int main () \{
  for (int i=0; i<512; i++) \{
    a[i] = b[i] + c[i];
  \}
\}

gcc -O3 -o ex2 ex2.c

00000000004003e0 <main>:
4003e0:31 c0                xor    %eax,%eax
4003e2:66 0f 1f 44 00 00    nopw   0x0(%rax,%rax,1)
4003e8:66 0f 6f 80 40 10 60 movdqa 0x601040(%rax),%xmm0
4003ef:00 
4003f0:48 83 c0 10          add    $0x10,%rax
4003f4:66 0f fe 80 30 18 60 paddd  0x601830(%rax),%xmm0
4003fb:00 
4003fc:0f 29 80 30 20 60 00 movaps %xmm0,0x602030(%rax)
400403:48 3d 00 08 00 00    cmp    $0x800,%rax
400409:75 dd                jne    4003e8 <main+0x8>
40040b:31 c0                xor    %eax,%eax
40040d:c3                   retq

@@

\begin{frame}[fragile]
\frametitle{Adding 512 64-bit integers - AVX2 superscalar instructions}
\begin{semiverbatim}
\tiny
int a[512] __attribute__((aligned(0x20)));
int b[512] __attribute__((aligned(0x20)));
int c[512] __attribute__((aligned(0x20)));

int main () \{
  for (int i=0; i<512; i++) \{
    a[i] = b[i] + c[i];
  \}
\}

gcc -march=native -O3 -o ex2 ex2.c

00000000004003e0 <main>:
4003e0:31 c0                xor    %eax,%eax
4003e2:66 0f 1f 44 00 00    nopw   0x0(%rax,%rax,1)
4003e8:c5 fd 6f 80 60 10 60 vmovdqa 0x601060(%rax),%ymm0
4003ef:00 
4003f0:c5 fd fe 80 60 18 60 vpaddd 0x601860(%rax),%ymm0,%ymm0
4003f7:00 
4003f8:48 83 c0 20          add    $0x20,%rax
4003fc:c5 fd 7f 80 40 20 60 vmovdqa %ymm0,0x602040(%rax)
400403:00 
400404:48 3d 00 08 00 00    cmp    $0x800,%rax
40040a:75 dc                jne    4003e8 <main+0x8>
40040c:c5 f8 77             vzeroupper 
40040f:31 c0                xor    %eax,%eax
400411:c3                   retq

@@

\begin{frame}[fragile]
\frametitle{Adding 512 64-bit integers - AVX512 superscalar instrs}
\begin{semiverbatim}
\tiny
int a[512] __attribute__((aligned(0x40)));
int b[512] __attribute__((aligned(0x40)));
int c[512] __attribute__((aligned(0x40)));

int main () \{
  for (int i=0; i<512; i++) \{
    a[i] = b[i] + c[i];
  \}
\}

gcc -mavx512vl -O3 -o ex2 ex2.c

00000000004003e0 <main>:
4003e0:31 c0                xor    %eax,%eax
4003e2:66 0f 1f 44 00 00    nopw   0x0(%rax,%rax,1)
4003e8:62 f1 fd 48 6f 80 80 vmovdqa64 0x601080(%rax),%zmm0
4003ef:10 60 00 
4003f2:48 83 c0 40          add    $0x40,%rax
4003f6:62 f1 7d 48 fe 80 40 vpaddd 0x601440(%rax),%zmm0,%zmm0
4003fd:14 60 00 
400400:62 f1 fd 48 7f 80 40 vmovdqa64 %zmm0,0x601840(%rax)
400407:18 60 00 
40040a:48 3d 00 08 00 00    cmp    $0x800,%rax
400410:75 d6                jne    4003e8 <main+0x8>
400412:31 c0                xor    %eax,%eax
400414:c3                   retq

baccala@ideapad-S510p:~$ ./ex2
Illegal instruction (core dumped)
baccala@ideapad-S510p:~$ ./ex2

@@

\begin{frame}[fragile]
\frametitle{Pointer Aliasing}
\begin{semiverbatim}
\tiny
void operation (int * a, int * b, int * c)
\{
  for (int i=0; i<16; i++) \{
    c[i] = a[i] + b[i];
  \}
\}
\end{semiverbatim}
\pause
\begin{semiverbatim}
\tiny
int v[18] = \{1, 1\};

@@

@@

@@

% \task - takes four arguments
% 1. name of task
% 2. location of task box's center point
% 3. "r" or "l", indicating if a set of tick marks representing
%    ports should drawn on the left or right side of the task box
%    "u" to make the task unlabled
% 4. contents of the task box - other nodes
%
% The tick marks will be named "#1 Port #" for future reference

\newcommand\task[4]{
\matrix (#1) [task, inner xsep=0] at (#2) { \node {\IfSubStr{#3}{u}{}{#1}}; \\ #4

@@

@@

\begin{tikzpicture}[transform canvas={scale=0.4, yshift=10em, xshift=-15em}]

@@

\draw [color=violet, thick, -triangle 90] (Task 1 left Port 5) -- (Task 1 right Port 5);
\draw [color=violet, thick] (Task 2 left Port 0) -- +(1,0) |- (Task 1 right Port 5);
\draw [color=violet, thick, -triangle 90] (Task 2 left Port 5) -- (Task 2 right Port 5);
\draw [color=violet, thick, triangle 90-] (Task 2 left Port -5) -- (Task 2 right Port -5);

\filldraw[fill=white] (Task 1 right Port 5) ++(-4,-0.5) rectangle +(1.8,1);
\filldraw[fill=white, pattern=vertical lines] (Task 1 right Port 5) ++(-4,-0.5) rectangle +(1.8,1);

\filldraw[fill=white] (Task 2 right Port 5) ++(-4,-0.5) rectangle +(1.8,1);
\filldraw[fill=white, pattern=vertical lines] (Task 2 right Port 5) ++(-4,-0.5) rectangle +(1.8,1);

\filldraw[fill=white] (Task 2 right Port -5) ++(-4,-0.5) rectangle +(1.8,1);
\filldraw[fill=white, pattern=vertical lines] (Task 2 right Port -5) ++(-4,-0.5) rectangle +(1.8,1);

@@

\begin{frame}

\frametitle{Send Rights}

\task{Task 1}{0,0}{r}{
% print two sendright nodes, one to fill the message1 box to the correct size,
% and another one to actually draw it, since message1 will overwrite the
% first sendright
\node (sendright) [draw, fill=orange] {Send 41};
\node (message1) [fit=(sendright), message] {Message 1};
\node (sendright) [draw, fill=orange] {Send 41};

\node (message3) at (0, -7em) [message] {Message 2};
}

\task{Task 2}{15 em,0}{l}{
\node (sendright2) [draw, fill=orange] {Send 9};
\node (message2) [fit=(sendright2), message] {Message 1};
\node (sendright2) [draw, fill=orange] {Send 9};

\node (message4) at (0, -7em) [message] {Message 2};
}

\draw [color=violet, thick, <-] (node cs:name=message3,angle=20) -- +(.3,0) |- (Task 1 Port -5) node [midway, anchor=south, color=black] {41} -- (Task 2 Port -5)
    -- +(.5,0) node [near end, anchor=south, color=black] {9} |- (node cs:name=message4, angle=160);

\begin{frame}

\frametitle{Send Rights}

\task{Task 1}{0,0}{r}{
% print two sendright nodes, one to fill the message1 box to the correct size,
% and another one to actually draw it, since message1 will overwrite the
% first sendright
\node (sendright) [draw, fill=orange] {Send 41};
\node (message1) [fit=(sendright), message] {Message 1};
\node (sendright) [draw, fill=orange] {Send 41};
}
\task{Task 2}{15 em,0}{l}{
\node (sendright2) [draw, fill=orange] {Send 9};
\node (message2) [fit=(sendright2), message] {Message 1};
\node (sendright2) [draw, fill=orange] {Send 9};
%\node (message2) [message] {Message 1};
% \node (message2a) [message] at (1 em, -2 em) {Message 2};
}
\task{Task 1 }{0,-11 em}{r}{
\node (message3) [message] {Message 2};
}
\task{Task 2 }{15 em,-11 em}{l}{
\node (message4) [message] {Message 2};
}

%\draw [color=violet, thick, ->] (node cs:name=message3,angle=0) -- +(.3,0) |- (Task 3 Port 0) node [midway, anchor=south, color=black] {18} -- +(1,0) |- (Task 2 Port 3)
%    -- +(.3,0) |- (node cs:name=message2a, angle=160);

\begin{frame}
\frametitle{Right Transfers}
\end{frame}

\begin{frame}
\frametitle{Memory Transfers}
\end{frame}

@@

@@

\begin{frame}
\frametitle{Mach Memory Management}
\end{frame}

@@

@@

>>>>

\usepackage{xcolor}
\usepackage{comment}
\usepackage{graphicx}

@@

\begin{frame}
\frametitle{Techniques to exploit parallelism}
\begin{itemize}
\item Instruction level parallelism
\begin{itemize}
\item Superscalar vector operations
\item Speculative execution
\item Co-processors
\end{itemize}
\item Thread level parallelism
\begin{itemize}
\item Shared memory
\item Caching
\item Synchronization
\end{itemize}
\item Node level parallelism
\begin{itemize}
\item No shared memory
\item High speed LAN
\item Clusters
\end{itemize}
\end{itemize}
\end{frame}

@@

int main () \{
  for (int i=0; i<512; i++) \{
    c[i] = a[i] + b[i];
  \}
\}

baccala@ideapad-S510p:~\$ gcc -O2 -fno-align-loops -o ex1 ex1.c

00000000004003e0 <main>:
4003e0:31 c0                xor    %eax,%eax
4003e2:8b 90 60 18 60 00    mov    0x601860(%rax),%edx
4003e8:03 90 60 10 60 00    add    0x601060(%rax),%edx
4003ee:48 83 c0 04          add    $0x4,%rax
4003f2:89 90 5c 20 60 00    mov    %edx,0x60205c(%rax)
4003f8:48 3d 00 08 00 00    cmp    $0x800,%rax
4003fe:75 e2                jne    4003e2 <main+0x2>
400400:31 c0                xor    %eax,%eax
400402:c3                   retq

@@

\begin{frame}[fragile]
\frametitle{Adding 512 64-bit integers - SSE superscalar instructions}
\begin{semiverbatim}
\tiny
int a[512];
int b[512];
int c[512];

int main () \{
  for (int i=0; i<512; i++) \{
    c[i] = a[i] + b[i];
  \}
\}

baccala@ideapad-S510p:~$ gcc -O3 -fno-align-loops -o ex1 ex1.c

00000000004003e0 <main>:
4003e0:31 c0                         xor    %eax,%eax
4003e2:66 0f 6f 80 60 10 60 00       movdqa 0x601060(%rax),%xmm0
4003ea:48 83 c0 10                   add    $0x10,%rax
4003ee:66 0f fe 80 50 18 60 00       paddd  0x601850(%rax),%xmm0
4003f6:0f 29 80 50 20 60 00          movaps %xmm0,0x602050(%rax)
4003fd:48 3d 00 08 00 00             cmp    $0x800,%rax
400403:75 dd                         jne    4003e2 <main+0x2>
400405:31 c0                         xor    %eax,%eax
400407:c3                            retq

@@

\begin{frame}[fragile]
\frametitle{Adding 512 64-bit integers - AVX2 superscalar instructions}
\begin{semiverbatim}
\tiny
int a[512] __attribute__((aligned(32)));
int b[512] __attribute__((aligned(32)));
int c[512] __attribute__((aligned(32)));

int main () \{
  for (int i=0; i<512; i++) \{
    c[i] = a[i] + b[i];
  \}
\}

baccala@ideapad-S510p:~$ gcc -march=native -O3 -o ex2 ex2.c

00000000004003e0 <main>:
4003e0:31 c0                         xor    %eax,%eax
4003e2:c5 fd 6f 80 60 10 60 00       vmovdqa 0x601060(%rax),%ymm0
4003ea:c5 fd fe 80 60 18 60 00       vpaddd 0x601860(%rax),%ymm0,%ymm0
4003f2:48 83 c0 20                   add    $0x20,%rax
4003f6:c5 fd 7f 80 40 20 60 00       vmovdqa %ymm0,0x602040(%rax)
4003fe:48 3d 00 08 00 00             cmp    $0x800,%rax
400404:75 dc                         jne    4003e2 <main+0x2>
400406:c5 f8 77                      vzeroupper 
400409:31 c0                         xor    %eax,%eax
40040b:c3                            retq

@@

\begin{frame}[fragile]
\frametitle{Adding 512 64-bit integers - AVX512 superscalar instrs}
\begin{semiverbatim}
\tiny
int a[512] __attribute__((aligned(64)));
int b[512] __attribute__((aligned(64)));
int c[512] __attribute__((aligned(64)));

int main () \{
  for (int i=0; i<512; i++) \{
    c[i] = a[i] + b[i];
  \}
\}

baccala@ideapad-S510p:~$ gcc -mavx512vl -O3 -o ex3 ex3.c

00000000004003e0 <main>:
4003e0:31 c0                         xor    %eax,%eax
4003e2:62 f1 fd 48 6f 80 80 10 60 00 vmovdqa64 0x601080(%rax),%zmm0
4003ec:48 83 c0 40                   add    $0x40,%rax
4003f0:62 f1 7d 48 fe 80 40 18 60 00 vpaddd 0x601840(%rax),%zmm0,%zmm0
4003fa:62 f1 fd 48 7f 80 40 20 60 00 vmovdqa64 %zmm0,0x602040(%rax)
400404:48 3d 00 08 00 00             cmp    $0x800,%rax
40040a:75 d6                         jne    4003e2 <main+0x2>
40040c:31 c0                         xor    %eax,%eax
40040e:c3                            retq

baccala@ideapad-S510p:~$ ./ex3
Illegal instruction (core dumped)
baccala@ideapad-S510p:~$ ./ex3

@@

\begin{frame}[fragile]
\frametitle{Pointer Aliasing}
\begin{semiverbatim}
\tiny
void operation (int * ap, int * bp, int * cp)
\{
  for (int i=0; i<16; i++) \{
    cp[i] = ap[i] + bp[i];
  \}
\}
\end{semiverbatim}
\pause
\begin{semiverbatim}
\tiny
int v[18] = \{1, 1\};

@@

\begin{frame}[fragile]
\frametitle{{\tt clDeviceQuery} on an Intel\textsuperscript{\textregistered} Core\texttrademark\ i5-4200U CPU}
\begin{semiverbatim}
\tiny
  CL_DEVICE_NAME:                       Intel(R) HD Graphics Haswell Ultrabook GT2 Mobile
  CL_DEVICE_VENDOR:                     Intel
  CL_DRIVER_VERSION:                    1.1.1
  CL_DEVICE_TYPE:                       CL_DEVICE_TYPE_GPU
  CL_DEVICE_MAX_COMPUTE_UNITS:          20
  CL_DEVICE_MAX_WORK_ITEM_DIMENSIONS:   3
  CL_DEVICE_MAX_WORK_ITEM_SIZES:        512 / 512 / 512 
  CL_DEVICE_MAX_WORK_GROUP_SIZE:        512
  CL_DEVICE_MAX_CLOCK_FREQUENCY:        1000 MHz
  CL_DEVICE_ADDRESS_BITS:               32
  CL_DEVICE_MAX_MEM_ALLOC_SIZE:         1024 MByte
  CL_DEVICE_GLOBAL_MEM_SIZE:            2048 MByte
  CL_DEVICE_ERROR_CORRECTION_SUPPORT:   no
  CL_DEVICE_LOCAL_MEM_TYPE:             global
  CL_DEVICE_LOCAL_MEM_SIZE:             64 KByte
  CL_DEVICE_MAX_CONSTANT_BUFFER_SIZE:   131072 KByte
  CL_DEVICE_QUEUE_PROPERTIES:           CL_QUEUE_PROFILING_ENABLE
  CL_DEVICE_IMAGE_SUPPORT:              1
  CL_DEVICE_MAX_READ_IMAGE_ARGS:        128
  CL_DEVICE_MAX_WRITE_IMAGE_ARGS:       8

  CL_DEVICE_IMAGE <dim>                 2D_MAX_WIDTH     8192
                                        2D_MAX_HEIGHT    8192
                                        3D_MAX_WIDTH     8192
                                        3D_MAX_HEIGHT    8192
                                        3D_MAX_DEPTH     2048

  CL_DEVICE_PREFERRED_VECTOR_WIDTH_<t>  CHAR 16, SHORT 8, INT 4, LONG 2, FLOAT 4, DOUBLE 0
\end{semiverbatim}
\end{frame}

\begin{frame}
\frametitle{Cluster Computing - History}
\begin{itemize}
\item Beowulf
\item Hadoop
\item Cluster Linux
\begin{itemize}
\item Kerrighed - Latest version is 3.0.0 which was released on 14 June 2010 and based on Linux kernel 2.6.30.
\item OpenSSI - last development release 2010; last stable release 2007
\end{itemize}
\end{itemize}
\end{frame}

@@

\tikzstyle{kernel} = [draw, fill=blue!20, rounded corners, minimum width=30em, minimum height=8em]

@@

% \task - takes four arguments
% 1. name of task (used to refer to it from tikz; maybe printed as a label as well)
% 2. location of task box's center point
% 3. "r" or "l", indicating if a set of tick marks representing
%    ports should drawn on the left or right side of the task box
%    "u" to label the task "Task" instead of labeling it with its name
%    "b" to make a "big" node (I wish I knew how to size the tick marks automatically)
% 4. contents of the task box - other nodes
%
% The tick marks will be named "#1 Port #" for future reference

\newcommand\task[4]{
\matrix (#1) [task, inner xsep=0] at (#2) { \node {\IfSubStr{#3}{u}{Task}{#1}}; \\ #4

@@

\newcommand\bigtask[4]{
\matrix (#1) [task, inner xsep=0] at (#2) { \node {\IfSubStr{#3}{u}{Task}{#1}}; \\ #4

\foreach \i in {-20,...,7}
{
\IfSubStr{#3}{l}{
  \draw[xshift=-5em, ultra thick] (0,0.2 * \i)  coordinate (#1 Port \i) -- (0.2,0.2 * \i);
}{
  \draw[xshift=-5em, ultra thick, color=blue!20] (0,0.2 * \i) -- (0.2,0.2 * \i);
}
\IfSubStr{#3}{r}{
  \draw[xshift=5em, ultra thick] (-0.2,0.2 * \i) -- (0,0.2 * \i) coordinate (#1 Port \i);
}{
  \draw[xshift=5em, ultra thick, color=blue!20] (-0.2,0.2 * \i) -- (0,0.2 * \i);
}
}

% draw a box with vertical lines to represent a message queue
% arg 1 is the name of a port to position it to the left of
% arg 2 is name to be given the box

\newcommand\queue[2]{
\filldraw (#1) +(-3.1,0) node[rectangle, name=#2, minimum width=1.8cm, minimum height=1cm, draw, pattern=vertical lines] {};
}

@@

\begin{tikzpicture}[transform canvas={scale=0.4, yshift=12em, xshift=-15em}]

@@

\queue{Task 1 right Port 5}{queue 1}
\queue{Task 2 right Port -5}{queue 2a}
\queue{Task 2 right Port 5}{queue 2b}
\queue{Task 3 right Port 0}{queue 3}
\queue{Task 4 right Port 0}{queue 4}

\draw [color=violet, thick, -triangle 90] (Task 1 left Port 5) -- (queue 1.west);
\draw [color=violet, thick, -triangle 90] (queue 1.east) -- (Task 1 right Port 5);
\draw [color=violet, thick] (Task 2 left Port 0) -- +(1,0) |- (queue 1.west);
\draw [color=violet, thick, -triangle 90] (Task 2 left Port -3) -- +(1,0) |- (queue 4.west);

\draw [color=violet, thick, -triangle 90] (Task 2 left Port 5) -- (queue 2b.west);
\draw [color=violet, thick, triangle 90-] (Task 2 left Port -5) -- (queue 2a.west);
\draw [color=violet, thick, -triangle 90] (queue 2b.east) -- (Task 2 right Port 5);
\draw [color=violet, thick, triangle 90-] (queue 2a.east) -- (Task 2 right Port -5);

\draw [color=violet, thick, -triangle 90] (Task 3 left Port 0) -- (queue 3.west);
\draw [color=violet, thick, -triangle 90] (queue 3.east) -- (Task 3 right Port 0);

\draw [color=violet, thick, -triangle 90] (Task 3 right Port 0) ++(-1,0) |- ++(-4,1) -- ++(0,-1);

\draw [color=violet, thick, -triangle 90] (queue 4.east) -- (Task 4 right Port 0);

\draw [color=violet, thick, -triangle 90] (Task 4 left Port 3) -- +(2,0) |- +(6,5.5) |- (queue 2a.east);

@@

\begin{frame}
\frametitle{Transfering a Send Right}
\begin{center}
\begin{tikzpicture}[node distance = .5cm]

\bigtask{Task 1}{0,0}{r}{
% print two sendright nodes, one to fill the message1 box to the correct size,
% and another one to actually draw it, since message1 will overwrite the
% first sendright
\only<2->{
\node (sendright) [draw, fill=orange] {Send 41};
\node (message1) [fit=(sendright), message] {Message 1};
\node (sendright) [draw, fill=orange] {Send 41};
}

\only<4->{\node (message3) at (0, -7em) [message] {Message 2};}
}

\bigtask{Task 2}{15 em,0}{l}{
\only<3->{
\node (sendright2) [draw, fill=orange] {Send 9};
\node (message2) [fit=(sendright2), message] {Message 1};
\node (sendright2) [draw, fill=orange] {Send 9};
}

\only<4->{\node (message4) at (0, -7em) [message] {Message 2};}
}

\only<2->{\node (port 41) [left of=Task 1 Port -10, anchor=south] {41};}

\only<2->{\draw [thick, color=red, dotted, ->] (node cs:name=sendright, angle=-20) -- (port 41);}

\node [left of=Task 1 Port 5, anchor=south] {27};
\node [right of=Task 2 Port 5, anchor=south] {52};
\only<1,2>{\draw [thick, -triangle 90] (Task 1 Port 5) -- (Task 2 Port 5);}

\only<3->{
\draw [color=violet, thick, -triangle 90] (node cs:name=message1,angle=20) -- +(.3,0) |- (Task 1 Port 5) -- (Task 2 Port 5)
    -- +(.5,0) |- (node cs:name=message2, angle=160);
}

\only<3->{\node [right of=Task 2 Port -10, anchor=south] {9};}

\only<3>{\draw [thick, triangle 90-] (Task 1 Port -10) -- (Task 2 Port -10);}

\only<4->{
\draw [color=violet, thick, triangle 90-] (node cs:name=message3,angle=20) -- +(.3,0) |- (Task 1 Port -10) -- (Task 2 Port -10)
    -- +(.5,0) |- (node cs:name=message4, angle=160);
}

\end{tikzpicture}
\end{center}
\end{frame}

\begin{frame}
\frametitle{Transfering a Receive Right}
\begin{center}
\begin{tikzpicture}[node distance = .5cm]

\bigtask{Task 1}{0,0}{r}{
% print two sendright nodes, one to fill the message1 box to the correct size,
% and another one to actually draw it, since message1 will overwrite the
% first sendright
\only<2->{
\node (sendright) [draw, fill=green] {Recv 41};
\node (message1) [fit=(sendright), message] {Message 1};
\node (sendright) [draw, fill=green] {Recv 41};
}

\only<4->{\node (message3) at (0, -7em) [message] {Message 2};}
}

\bigtask{Task 2}{15 em,0}{l}{
\only<3->{
\node (sendright2) [draw, fill=green] {Recv 9};
\node (message2) [fit=(sendright2), message] {Message 1};
\node (sendright2) [draw, fill=green] {Recv 9};
}

\only<4->{\node (message4) at (0, -7em) [message] {Message 2};}
}

\only<2->{\node (port 41) [left of=Task 1 Port -10, anchor=south] {41};}

\only<2->{\draw [thick, color=red, dotted, ->] (node cs:name=sendright, angle=-20) -- (port 41);}

\node [left of=Task 1 Port 5, anchor=south] {27};
\node [right of=Task 2 Port 5, anchor=south] {52};
\only<1,2>{\draw [thick, -triangle 90] (Task 1 Port 5) -- (Task 2 Port 5);}

\only<3->{
\draw [color=violet, thick, -triangle 90] (node cs:name=message1,angle=20) -- +(.3,0) |- (Task 1 Port 5) -- (Task 2 Port 5)
    -- +(.5,0) |- (node cs:name=message2, angle=160);
}

\only<3->{\node [right of=Task 2 Port -10, anchor=south] {9};}

\only<3>{\draw [thick, -triangle 90] (Task 1 Port -10) -- (Task 2 Port -10);}

\only<4->{
\draw [color=violet, thick, -triangle 90] (node cs:name=message3,angle=20) -- +(.3,0) |- (Task 1 Port -10) -- (Task 2 Port -10)
    -- +(.5,0) |- (node cs:name=message4, angle=160);
}

\end{tikzpicture}
\end{center}
\end{frame}

\begin{frame}
\frametitle{Memory Transfers}
\begin{center}

\begin{tikzpicture}[node distance = .5cm]

\bigtask{Task 1}{0,0}{r}{
% print two sendright nodes, one to fill the message1 box to the correct size,
% and another one to actually draw it, since message1 will overwrite the
% first sendright
\node (message1) [message] {Message};
\only<1>{
\node (memory1) [below=1em, fill=orange!40, minimum size=2cm, dashed, thick, draw] at (message1.south) {Memory};
\draw [thick, color=red, dotted, -] (message1) -- (memory1);
}
}

\bigtask{Task 2}{15 em,0}{l}{
\only<2->{
\node (message2) [message] {Message};
\node (memory2) [below=1em, fill=orange!40, minimum size=2cm, dashed, thick, draw] at (message2.south) {Memory};
\draw [thick, color=red, dotted, -] (message2) -- (memory2);
}
}

%\node [left of=Task 1 Port 5, anchor=south] {27};
%\node [right of=Task 2 Port 5, anchor=south] {52};

%\only<1>{\draw [thick, -triangle 90] (Task 1 Port 5) -- (Task 2 Port 5);}

\only<2>{
\draw [color=violet, thick, -triangle 90] (node cs:name=message1,angle=20) -- +(.3,0) |- (Task 1 Port 5) -- (Task 2 Port 5)
    -- +(.5,0) |- (node cs:name=message2, angle=160);
}
\end{tikzpicture}
\end{center}
\end{frame}

@@

\begin{frame}[fragile]
\frametitle{{\tt rpctrace} of {\tt ls} (9 of 144 RPCs)}
\begin{semiverbatim}
\tiny
task135(pid9403)->task_get_special_port (4) = 0    143(pid9403)

  143(pid9403)->exec_startup_get_info () = 0 134526367 134512692 288 188416 16777216 0 "ls\\0"
        "LS_COLORS=rs=0:di=01;34:ln=01;36:mh=00:pi=40;33:so=01;35:do=01;35:bd=40;33;01:cd"
        \{  140(pid9403)   145(pid9403)   146(pid9403)   147(pid9403)\}
        \{  148(pid9403)   149(pid9403)   150(pid9403)   151(pid9403)   152(pid9403) (null)\}
        \{18 0 0 0 0\}

/* Standard port assignments for file_exec_file_name and exec_* */
enum
  \{
    INIT_PORT_CWDIR,
    INIT_PORT_CRDIR,
    INIT_PORT_AUTH,
    INIT_PORT_PROC,
    INIT_PORT_CTTYID,
    INIT_PORT_BOOTSTRAP,
  \};

  145(pid9403)->term_open_ctty (9403 9402) = 0    138(pid9403)

  148(pid9403)->dir_lookup ("./" 2097161 0) = 0 1 "    166(pid9403)

  166(pid9403)->dir_readdir (0 -1 0) = 0 "\\x01\\x9c\\x01\\0\\0\\0\\0\\0\\x10\\0\\0\\x01.\\0\\0\\0P\\xd7\\0\\0\\0\\0\\0\\0\\x10\\0\\0\\x02..\\0\\0\\x03\\x9c\\x01\\0\\0\\0\\0\\0\\x14\\0\\0\\x04.git\\0\\0\\0\\0\\x90_\\a\\0\\0\\0\\0\\0\\x18\\0\\0\\blibpager\\0\\0\\0\\0c\\x9a\\x01\\0" 10

  166(pid9403)->dir_readdir (10 -1 0) = 0 " 0

  145(pid9403)->io_stat () = 0 {14 969 0 0 0 1458877724 0 8397184 1 0 5 0 0 1515530623 0 1515530623 0 1515530623 0 512 8 0 0 0 0 0 1103856 1038 1163584 1039 0 0}

  138(pid9403)->io_write ("junk  libpager machMessage.h  netmsg  papers  patches README.md\\n" -1) = 0 64

task135(pid9403)->task_terminate () = 0 
\end{semiverbatim}
\end{frame}

@@

\begin{frame}
\frametitle{Mach Memory Management}
\begin{center}
\begin{tikzpicture}

\task{Filesystem Server}{0,0}{}{
\node (memory1) [above left=-0.5, fill=white, minimum size=1.5cm, dashed, thick, draw] {};
}
\task{ls}{15 em,0}{}{
%\node (memory1) [above right=-0.5, fill=orange!40, minimum size=1.5cm, dashed, thick, draw] {};
\node (memory2) [above right=-0.5, fill=white, minimum size=1.5cm, dashed, thick, draw] {};
\node [below] at (memory2.north) {\tiny addr};
}

\node (kernel) [kernel] at (7.5 em, -10 em) {};
\node (kmemory) [fill=white, minimum size=1.5cm, dashed, thick, draw] at (7.5 em, -10 em) {};
\node [below] at (kernel.north) {Kernel};

\draw [color=red, <-, thick] (2 em,1) -- node [midway, above] {io\_map} (13 em,1);
\draw [color=red, ->, thick, dotted] (2 em,.9) -- node [midway, below] {\tiny port 153} (13 em,.9);

\draw [color=red, ->, thick] (15 em,-3 em) -- node [midway, left] {vm\_map} (15 em,- 8 em);
\draw [color=red, <-, thick, dotted] (15.2 em,-3 em) -- node [midway, right] {\tiny addr} (15.2 em,- 8 em);

\draw [color=red, ->, thick] (-1.5 em,-3 em) -- node [midway, right] {m\_o\_ready} (-1.5 em,- 8 em);
\draw [color=red, <-, thick] (-2 em,-3 em) -- node [midway, left] {m\_o\_init} (-2 em,- 8 em);

\end{tikzpicture}
\end{center}
\end{frame}

@@

\begin{frame}[fragile]
\frametitle{{\tt rpctrace} of {\tt ls} (mapping the C library)}
\begin{semiverbatim}
\tiny
  149(pid9403)->dir_lookup ("lib/i386-gnu/libc.so.0.3" 1 0) = 0 1 ""    155(pid9403)

  155(pid9403)->io_read (-1 512) = 0 "\\x7fELF\\x01\\x01\\x01\\0\\0\\0\\0\\0\\0\\0\\0\\0\\x03\\0\\x03\\0\\x01\\0\\0\\0pE\\x05\\04\\0\\0\\0\\xfc\\x19''\\0\\0\\0\\0\\04\\0 \\0\\n\\0(\\0x\\0w\\0\\x06\\0\\0\\04\\0\\0\\04\\0\\0\\04\\0\\0\\0@\\x01\\0\\0@\\x01\\0\\0\\x05\\0\\0\\0"

  155(pid9403)->io_stat () = 0 {23 5 0 158020 0 1513754748 0 33261 1 0 0 2239676 0 1515530623 0 1513435063 0 1513747658 0 8192 4384 0 0 0 0 0 0 0 0 0 0 0}

  155(pid9403)->io_map () = 0    153(pid9403)  (null)

task135(pid9403)->vm_map (0 2245448 0 1    153(pid9403) 0 32 5 7 1) = 0 17039360

\begin{frame}
\frametitle{The {\tt netmsg} Server}
\end{frame}

@@

\begin{frame}
\frametitle{What kind of bugs have I found?}
\begin{itemize}
\item failure to update pagemap entry
\item FLUSHING flag
\item DEAD NAME notifications
\item condition variable
\item weak refs
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Work Needed / Help Wanted}
\begin{itemize}
\item 64-bit user space
\item Multiprocessor support
\item Rewrite of {\tt netmsg}
\end{itemize}
\end{frame}
